Variation in site entropy explains differences in structure-sequence relationships of proteins

### 2015-06-01 manuscript/manuscript.tex
Adding a discussion

\section{Discussion}
\label{sec:dcr}
{\color{red}SOME DISCUSSION.....} \\

>>>>

\section{Discussion}
\label{sec:dcr}
Structure imposes constraints on the the evolution of proteins. Therefore structural predictors can be used to predict the variability at sites. Two main determinants of variability at protein sits are solvent accessibility and local packing density. Sites that are on surface of the protein tend to have higher solvent accessibility and exhibit more site variability. Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability. However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins. Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the main determinant. In a study done on monomeric enzyme proteins, \cite{Yehetal2014} reported that local packing density, as measured by Contact Number and Weighted Contact Number, is the strongest determinant of site variability. \cite{Shahmoradietal2014} found that which a group of virus proteins, solvent accessibility is more predictive of site-wise variability.

\indent Here we performed examined the relationship between site variability and the strength of structure-sequence relationships. We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average. However, when looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets. Overall the enzyme proteins have a higher site variability than the virus proteins. When predicting the structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variations. It appears that site variability and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships. Additionally the reason that the enzyme proteins in \cite{Yehetal2014} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}. Moreover, since the variance is entropy is is a more significant predictor of correlation's in local packing density, an increase level of variance in site variability will have a larger effect on the local packing density correlations. This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of /\cite{Yehetal2014}. Finally, The variance in evolutionary rate among sites of a protein does not have an effect on the correlation magnitude. This suggests that the absolute variability at sites is a better predictor of whether structural determinants can be used to predict variability at sites  as opposed to evolutionary rate. Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
